--- 
title: "Biostatistics"
author: "David Coffey"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [Citations.bib]
biblio-style: apalike
link-citations: yes
github-repo: davidcoffey/Biostatistics
description: "Principles of biostatistics and how to apply them using the R programming language."
---
--- 
title: "Biostatistics"
author: "David Coffey"
date: "`r Sys.Date()`"
site: bookdown::bookdown_site
output: bookdown::gitbook
documentclass: book
bibliography: [Citations.bib]
biblio-style: apalike
link-citations: yes
github-repo: davidcoffey/Biostatistics
description: "Principles of biostatistics and how to apply them using the R programming language."
---

# General overview

## Introduction
This books provides a consise overview of biostatistics and its applications using the R programming language.  The textbook *Fundamentals of Biostatitics* [@Rosner2016] was used extensivity in the preparation of this book.

## Example dataset
Examples of R functions are performed on a dataset of patients with newly diagnosed multiple myeloma.  This dataset  contains a variety of categorical and continuous variables. A description of the variables are shown below.

```{r echo=FALSE}
library(knitr)
library(kableExtra)
load("docs/Data-dictionary.Rda")
kable(dictionary)
```

<!--chapter:end:index.Rmd-->


# Descriptive statistics

Placeholder


## Measures of location
### Arithmetic mean
### Median
### Mode
### Geometric mean
## Measures of spread
### Range
### Percentile
### Variance
### Standard deviation
### Standard error of the mean
### Coefficient of varaince
## Visualizing descriptive statistics
### Single variable
#### Stem and leaf plot
#### Frequency distribution table
#### Dot plot
#### Histogram
#### Density plot
### Muliple variables
#### Summary table
#### Bar graph
#### Box plot
#### Violin plot
#### Jitter plot

<!--chapter:end:DescriptiveStatistics.Rmd-->


# Probability

Placeholder


## Table of confusion
## Sensitivity
## Specificity
## Postitive predictive value
## Negative predictive value
## ROC curve
## Baysian inference

<!--chapter:end:Probability.Rmd-->

# Normal distribution

## Probability distributions
A probability distribution is a mathematical function that provides the probabilities of occurrence of different possible outcomes.  Probability distributions are generally divided into two classes. A **discrete probability distribution** applies to discrete random variables such as the probability of being alive or dead as in the binomial distribution or the probability of a given number of deaths occurring in a fixed time interval as in the Poisson distribution.  A **continuous probability distribution** applies to continuous random variables such as the measure of hematocrit on any given day.  The most common continuous probability distribution is the normal or Gaussian distribution and is defined by the following probability-density function:
$$f(x)=\frac{1}{\sqrt{2\pi\sigma}}e^{-\frac{1}{2\sigma^{2}}(x-\mu)^{2}}$$
where ${x}$ is a continuous random variable, ${\sigma^{2}}$ is the variance, and ${\mu}$ is the mean.  Properties of the normal distribution include:

* The curve is symmetric about ${\mu}$
* The entire shape of the normal distribution is determined by the two parameters ${\mu}$ and ${\sigma^{2}}$
* The area under the curve between any two points ${a}$ and ${b}$ is equal to the probability that the random variable ${x}$ falls between ${a}$ and ${b}$
* The area under the entire normal density function is always equal to 1
* The probability that the random variable ${x}$ falls between ±1 standard deviation is approximately 68%, ±2 standard deviations is 95%, and ±2.5 standard deviations is 99%.

```{r, comment = NA}
# Calculate the probabilities of a normally distributed random variable
library(ggplot2)
p = dnorm(1:10, mean = 5, sd = 1)
p
ggplot(data.frame(Probability = p, x = 1:10), aes(x = x, y = Probability)) +
  geom_density() +
  geom_bar(stat = "identity")
```

<!--chapter:end:NormalDistribution.Rmd-->

`r if (knitr:::is_html_output()) '
# References {-}
'`

<!--chapter:end:References.Rmd-->

